Kristi Meeting

### 2024-09-27 KristiPresentationData.Rmd
Dina Folder Update

---
title: "OXPHOS KEGG Genes"
author: "John Santiago"
date: "2024-06-22"
output: html_document
---

@@

```{r echo = F,include = F}

@@

gbbOE = read.csv("/Users/johncsantiago/Documents/GitHub/Datasets/gbbOE_CountTable.csv", row.names = 1)
gbbOE.groups = read.csv("/Users/johncsantiago/Documents/GitHub/Datasets/gbbOE_Metadata.csv", row.names = 1)
gbbOE = gbbOE[,row.names(gbbOE.groups)]
gbbOE.groups$group = paste0(gbbOE.groups$genotype, "_",gbbOE.groups$stage)

##normalize data
countdata=gbbOE
x <- countdata
gr <- factor(gbbOE.groups$group, levels=c(unique(gbbOE.groups$group)))
y <- DGEList(counts=x,group=gr)
keep <- filterByExpr(y)
y <- y[keep,,keep.lib.sizes=FALSE]
z <- calcNormFactors(y, method = "TMM")
##normalized cpm
gbbOE.cpmdata=cpm(z)

gbbOEmean = gbbOE.cpmdata[,1:6]
colnames(gbbOEmean) = unique(gr)
i=1
while(i<=6){
  temp= gbbOE.cpmdata[,row.names(gbbOE.groups[gbbOE.groups$group == colnames(gbbOEmean)[i],])]
  gbbOEmean[,i] = apply(temp, 1, mean)
  i=i+1
}

design<-model.matrix(~0 + gr)
colnames(design) = unique(gr)
z = estimateDisp(z, design, robust = T)

fit <- glmQLFit(z, design)
plotQLDisp(fit)

compare = makeContrasts(G85R_LL3-LoxP_LL3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.LL3=DEout
sGRxWT.LL3=DEout[DEout$FDR<.05,]

compare = makeContrasts(G85R_ML3-LoxP_ML3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.ML3=DEout
sGRxWT.ML3=DEout[DEout$FDR<.05,]

compare = makeContrasts(G85R_LL3-G85R_ML3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.LL3xML3=DEout
sGR.LL3xML3=DEout[DEout$FDR<.05,]

compare = makeContrasts(LoxP_LL3-LoxP_ML3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.LL3xML3=DEout
sWT.LL3xML3=DEout[DEout$FDR<.05,]

compare = makeContrasts(G85R_gbb_LL3-G85R_gbb_ML3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
G85R_gbb.LL3xML3=DEout
sG85R_gbb.LL3xML3=DEout[DEout$FDR<.05,]

gbbKO = read.csv("/Users/johncsantiago/Documents/GitHub/Datasets/gbbKO_CountTable.csv", row.names = 1)
gbbKO.groups = read.csv("/Users/johncsantiago/Documents/GitHub/Datasets/gbbKO_Metadata.csv", row.names = 1)

##normalize data
countdata=gbbKO
x <- countdata
gr <- factor(gbbKO.groups$Genotype, levels=c(unique(gbbKO.groups$Genotype)))
y <- DGEList(counts=x,group=gr)
keep <- filterByExpr(y)
y <- y[keep,,keep.lib.sizes=FALSE]
z <- calcNormFactors(y, method = "TMM")
##normalized cpm
gbbKO.cpmdata=cpm(z)

gbbKOmean = gbbKO.cpmdata[,1:2]
colnames(gbbKOmean) = c('WT', 'KO')
i=1
while(i<=2){
  temp= gbbKO.cpmdata[,row.names(gbbKO.groups)[gbbKO.groups$Genotype == colnames(gbbKOmean)[i]]]
  gbbKOmean[,i] = apply(temp, 1, mean)
  i=i+1
}

design<-model.matrix(~0 + gr)
colnames(design) = unique(gr)
z = estimateDisp(z, design, robust = T)

fit <- glmQLFit(z, design)
plotQLDisp(fit)

compare = makeContrasts(KO-WT, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
KOxWT=DEout
sKOxWT=DEout[DEout$FDR<.05,]

@@

hmdata = gbbOEmean[intersect(row.names(gbbOEmean),geneset),]
gbbOE.order = c("LoxP_ML3", "LoxP_LL3",
          "G85R_ML3", "G85R_LL3",
          "G85R_gbb_ML3", "G85R_gbb_LL3")

hmdata = (hmdata[!is.na((GeneIDKey[row.names(hmdata), 'Symbol'])),gbbOE.order])
row.names(hmdata) = GeneIDKey[row.names(hmdata), 'Symbol']

hmdata = gbbKO.cpmdata[intersect(row.names(gbbKOmean),geneset), row.names(gbbKO.groups)]
colnames(hmdata) = c("WT1", "WT2", "WT3", "KO1", "KO2", "KO3")

hmdata = (hmdata[!is.na((GeneIDKey[row.names(hmdata), 'Symbol'])),])
row.names(hmdata) = GeneIDKey[row.names(hmdata), 'Symbol']

```{r warning=F,  echo = F, include = F}

##colnames(G85R.FC) = "GRF.CxDf"  "GRF.CxOE"  "WTF.CxDf" "WTF.CxOE"  "GRxWT.FC"  "GRxWT.FDf" "GRxWT.FOE" "GRM.CxDf"  "GRM.CxOE" "WTM.CxDf"  "WTM.CxOE"  "GRxWT.MC" "GRxWT.MDf" "GRxWT.MOE" "GR.FxM" "WT.FxM"

##colnames(G85R.metabFC) = "KEGG" "GRxWT.C" "GRxWT.Df" "GRxWT.OE" "GR.CxDf"  "WT.CxDf" "GR.CxOE"  "WT.CxOE"

FC.data = G85R.FC[,c("WTF.CxDf", "WTM.CxDf")]
metab.data = setNames(G85R.metabFC[,"WT.CxDf"], G85R.metabFC$KEGG)
out.id = "WT_ControlxDeficient_FandM_Genes with WT_ControlxDeficient_Metab"

PathwayID = "00190"
sigs = row.names(FC.data)

##File name

OutputDir = "/Users/johncsantiago/Documents/GitHub/WhartonLab/G85R OXPHOS KEGG Pathways/"
setwd(OutputDir)
img.file = paste0(OutputDir, "dme",PathwayID, ".", out.id, ".multi.png")

@@

  #knitr::include_graphics(img.file)

@@

```{r warning=F,  echo = F, include = F}

##EdgeR Comparisons  = GRxWT.ML3 GR.LL3xML3 WT.LL3xML3 G85R_gbb.LL3xML3

FC.data = data.frame(FC1 = GR.LL3xML3[row.names(GRxWT.ML3), "logFC"],
                     FC2 = G85R_gbb.LL3xML3[row.names(GRxWT.ML3), "logFC"])
row.names(FC.data) = row.names(GRxWT.ML3)

out.id = "G85RandG85R.gbbOE_LL3xML3"

PathwayID = "00190"
sigs = row.names(FC.data)

##File name

OutputDir = "/Users/johncsantiago/Documents/GitHub/WhartonLab/G85R OXPHOS KEGG Pathways/"
setwd(OutputDir)
img.file = paste0(OutputDir, "dme",PathwayID, ".", out.id, ".multi.png")

    deg.data = as.matrix(FC.data)
    row.names(deg.data) = GeneIDKey[row.names(FC.data), "ensembl"]

  pv.out <- pathview(gene.data =  deg.data,
                     pathway.id = PathwayID,
                     species = "dme",
                     kegg.native = T,
                     limit=list(gene=c(-2,2)),
                     node.sum="max.abs",
                     low="royalblue",mid = "grey",high="firebrick",
                     out.suffix = out.id,
                     bins=list(genes=30),
                     plot.col.key = T,
                     match.data = F)

  #knitr::include_graphics(img.file)

@@

```{r warning=F,  echo = F, include = F}

##EdgeR Comparisons  = KOxWT

FC.data = setNames(KOxWT$logFC, row.names(KOxWT))

out.id = "gbbKOxgbbWT"

PathwayID = "00190"
sigs = names(FC.data)

##File name

OutputDir = "/Users/johncsantiago/Documents/GitHub/WhartonLab/G85R OXPHOS KEGG Pathways/"
setwd(OutputDir)
img.file = paste0(OutputDir, "dme",PathwayID, ".", out.id, ".png")

    deg.data = as.vector(FC.data)
    names(deg.data) = GeneIDKey[names(FC.data), "ensembl"]

  pv.out <- pathview(gene.data =  deg.data,
                     pathway.id = PathwayID,
                     species = "dme",
                     kegg.native = T,
                     limit=list(gene=c(-2,2)),
                     node.sum="max.abs",
                     low="royalblue",mid = "grey",high="firebrick",
                     out.suffix = out.id,
                     bins=list(genes=30),
                     plot.col.key = T,
                     match.data = F)

  #knitr::include_graphics(img.file)

@@

>>>>

---
title: "Kristi Meeting"
author: "John Santiago"
date: "2024-06-18"
output: html_document
---

@@

```{r echo = F}

@@

@@

```{r warning=F,  echo = F}

FC.data = G85R.FC[,c("GRF.CxOE", "WTF.CxOE")]
metab.data = setNames(G85R.metabFC[,"GR.CxOE"], G85R.metabFC$KEGG)
PathwayID = "00190"
sigs = row.names(FC.data)

##File name
out.id = "G85RandWT_CxOE_Genes with GR.CxOE_Metab"
OutputDir = "/Users/johncsantiago/Documents/GitHub/WhartonLab/G85R OXPHOS KEGG Pathways/"
setwd(OutputDir)
img.file = paste0(OutputDir, "dme",PathwayID, ".", out.id, ".multi.png")

@@

  knitr::include_graphics(img.file)

@@

```{r echo = F}

heatmaply(G85R.metabmean[c('nicotinamide', 'nicotinic acid', "NAD+", "NADH"),
                         c('WT', 'TktDfWT', 'TktOEWT', 'GR', 'TktDfGR', 'TktOEGR')],
          scale = 'row',
          Colv = F,
                 custom_hovertext = signif(G85R.metabmean[c('nicotinamide', 'nicotinic acid', "NAD+", "NADH"),
                         c('WT', 'TktDfWT', 'TktOEWT', 'GR', 'TktDfGR', 'TktOEGR')], 3))

@@

```{r echo = F}

PPP = names(kegg[grep('00030', kegg)])
PPP = row.names(GeneIDKey[GeneIDKey$Symbol %in% PPP, ])

geneset = PPP

hmdata = A4V.meancpm[geneset,]
row.names(hmdata) = GeneIDKey[geneset, 'Symbol']
A4V.order = c("S3FH", "S9FH", "S40FH", "A3FH", "A9FH", "A40FH",
              "S3FT", "S9FT", "S40FT", "A3FT", "A9FT", "A40FT",
              "S3FA", "S9FA", "S40FA", "A3FA", "A9FA", "A40FA",
              "S3MH", "S9MH", "A3MH", "A9MH",
              "S3MT", "S9MT", "A3MT", "A9MT",
              "S3MA", "S9MA", "A3MA", "A9MA")

hmdata = na.omit(hmdata[,A4V.order])

heatmaply(hmdata,
          scale = 'row',
          Colv = F)

@@

```{r echo = F}

hmdata = G85R.meancpm[geneset,]
row.names(hmdata) = GeneIDKey[geneset, 'Symbol']
order = c("WTFC", "WTFDf", "WTFOE",
          "GRFC", "GRFDf", "GRFOE",
          "WTMC", "WTMDf", "WTMOE",
          "GRMC", "GRMDf", "GRMOE")

hmdata = na.omit(hmdata[,order])

heatmaply(hmdata,
          scale = 'row',
          Colv = F)

```{r warning=F,  echo = F}

FC = G85R.FC[,c("GRxWT.FC", "GRxWT.MC")]
metab.data = setNames(G85R.metabFC[,"GRxWT.C"], G85R.metabFC$KEGG)
PathwayID = "00030"
use = 'both'

  ##File name prefix
  out.id = paste("G85RvsWT_FemaleandMale")

  ##table with the FC data you want to use for coloring with FBgn for row names 
  ##if only analyzing 1 condition, use setNames to make a named vector
  ##A4V.FC or G85R.FC

  FC.data =  FC

  ##all or a subset of row.names used in FC.data
  usegenes = row.names(FC)

  #Any column name from G85R.metabFC or 'none'
  metab.data = metab.data

  if(use == 'primary'){

    ##File name prefix
    out.id = comparison

    ##table with the FC data you want to use for coloring with FBgn for row names 
    ##if only analyzing 1 condition, use setNames to make a named vector
    ##A4V.FC or G85R.FC

    FC.data =  FC[,1]

    #Any column name from G85R.metabFC or 'none'
    metab.data = metab.data
  }

  ##KEGG Mapping Function
  if(length(intersect(list.files(), "KEGG_Image_Files")) == 0){
    dir.create(paste0(dirname(rstudioapi::getSourceEditorContext()$path), "/KEGG_Image_Files/"))
  }
  Home = dirname(rstudioapi::getSourceEditorContext()$path)
  OutputDir = paste0(Home, "/KEGG_Image_Files/")
  setwd(OutputDir)
  sigs = usegenes
  ##sigs = names(FDR.data[FDR.data <= .05])

  #if(length(G85R.metabFC[metab.id, names(metab.data)]) > 1 ){
   # metab.data = setNames(log2(G85R.metabFC[metab.id, metab.data]), names(metab.id))
  #}

  pv.out <- pathview(gene.data =  deg.data,
                     cpd.data = metab.data,
                     pathway.id = PathwayID,
                     species = "dme",
                     kegg.native = T,
                     limit=list(gene=c(-2,2)),
                     node.sum="max.abs",
                     low="royalblue",mid = "grey",high="firebrick",
                     out.suffix = out.id,
                     bins=list(genes=30),
                     plot.col.key = T,
                     match.data = F)
  setwd(Home)

  if(class(FC.data)=="numeric"){
    img.file = paste0(OutputDir, "dme",PathwayID, ".", out.id, ".png")
  }

  if(class(FC.data)=="data.frame"){
    img.file = paste0(OutputDir, "dme",PathwayID, ".", out.id, ".multi.png")
  }
  knitr::include_graphics(img.file)

```{r warning=F,  echo = F}

FC = G85R.FC[,c("GRF.CxDf", "WTF.CxDf")]
metab.data = setNames(G85R.metabFC[,"GR.CxDf"], G85R.metabFC$KEGG)
PathwayID = "00030"
use = 'both'

  ##File name prefix
  out.id = paste("G85RvsWT_FemaleandMale")

  ##table with the FC data you want to use for coloring with FBgn for row names 
  ##if only analyzing 1 condition, use setNames to make a named vector
  ##A4V.FC or G85R.FC

  FC.data =  FC

  ##all or a subset of row.names used in FC.data
  usegenes = row.names(FC)

  #Any column name from G85R.metabFC or 'none'
  metab.data = metab.data

  if(use == 'primary'){

    ##File name prefix
    out.id = comparison

    ##table with the FC data you want to use for coloring with FBgn for row names 
    ##if only analyzing 1 condition, use setNames to make a named vector
    ##A4V.FC or G85R.FC

    FC.data =  FC[,1]

    #Any column name from G85R.metabFC or 'none'
    metab.data = metab.data
  }

  ##KEGG Mapping Function
  if(length(intersect(list.files(), "KEGG_Image_Files")) == 0){
    dir.create(paste0(dirname(rstudioapi::getSourceEditorContext()$path), "/KEGG_Image_Files/"))
  }
  Home = dirname(rstudioapi::getSourceEditorContext()$path)
  OutputDir = paste0(Home, "/KEGG_Image_Files/")
  setwd(OutputDir)
  sigs = usegenes
  ##sigs = names(FDR.data[FDR.data <= .05])

  #if(length(G85R.metabFC[metab.id, names(metab.data)]) > 1 ){
   # metab.data = setNames(log2(G85R.metabFC[metab.id, metab.data]), names(metab.id))
  #}

  pv.out <- pathview(gene.data =  deg.data,
                     cpd.data = metab.data,
                     pathway.id = PathwayID,
                     species = "dme",
                     kegg.native = T,
                     limit=list(gene=c(-2,2)),
                     node.sum="max.abs",
                     low="royalblue",mid = "grey",high="firebrick",
                     out.suffix = out.id,
                     bins=list(genes=30),
                     plot.col.key = T,
                     match.data = F)
  setwd(Home)

  if(class(FC.data)=="numeric"){
    img.file = paste0(OutputDir, "dme",PathwayID, ".", out.id, ".png")
  }

  if(class(FC.data)=="data.frame"){
    img.file = paste0(OutputDir, "dme",PathwayID, ".", out.id, ".multi.png")
  }
  knitr::include_graphics(img.file)

```{r warning=F,  echo = F}

FC = G85R.FC[,c("GRF.CxOE", "WTF.CxOE")]
metab.data = setNames(G85R.metabFC[,"GR.CxOE"], G85R.metabFC$KEGG)

PathwayID = "00030"
use = 'both'

  ##File name prefix
  out.id = paste("G85RvsWT_FemaleandMale")

  ##table with the FC data you want to use for coloring with FBgn for row names 
  ##if only analyzing 1 condition, use setNames to make a named vector
  ##A4V.FC or G85R.FC

  FC.data =  FC

  ##all or a subset of row.names used in FC.data
  usegenes = row.names(FC)

  #Any column name from G85R.metabFC or 'none'
  metab.data = metab.data

  if(use == 'primary'){

    ##File name prefix
    out.id = comparison

    ##table with the FC data you want to use for coloring with FBgn for row names 
    ##if only analyzing 1 condition, use setNames to make a named vector
    ##A4V.FC or G85R.FC

    FC.data =  FC[,1]

    #Any column name from G85R.metabFC or 'none'
    metab.data = metab.data
  }

  ##KEGG Mapping Function
  if(length(intersect(list.files(), "KEGG_Image_Files")) == 0){
    dir.create(paste0(dirname(rstudioapi::getSourceEditorContext()$path), "/KEGG_Image_Files/"))
  }
  Home = dirname(rstudioapi::getSourceEditorContext()$path)
  OutputDir = paste0(Home, "/KEGG_Image_Files/")
  setwd(OutputDir)
  sigs = usegenes
  ##sigs = names(FDR.data[FDR.data <= .05])

  #if(length(G85R.metabFC[metab.id, names(metab.data)]) > 1 ){
   # metab.data = setNames(log2(G85R.metabFC[metab.id, metab.data]), names(metab.id))
  #}

  pv.out <- pathview(gene.data =  deg.data,
                     cpd.data = metab.data,
                     pathway.id = PathwayID,
                     species = "dme",
                     kegg.native = T,
                     limit=list(gene=c(-2,2)),
                     node.sum="max.abs",
                     low="royalblue",mid = "grey",high="firebrick",
                     out.suffix = out.id,
                     bins=list(genes=30),
                     plot.col.key = T,
                     match.data = F)
  setwd(Home)

  if(class(FC.data)=="numeric"){
    img.file = paste0(OutputDir, "dme",PathwayID, ".", out.id, ".png")
  }

  if(class(FC.data)=="data.frame"){
    img.file = paste0(OutputDir, "dme",PathwayID, ".", out.id, ".multi.png")
  }
  knitr::include_graphics(img.file)
